---
title: "Data pipelines"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
write_csv(iris, "iris.csv")
```

## read

```{r}
read_csv("iris.csv")
iris_raw = read_csv("iris.csv")
```
## you can also read and write in excel
```{r}
library(readxl)
library(writexl)
write_xlsx(iris, "iris.xlsx")
read_excel("iris.xlsx")
```

##Exploratory analysis
Data explorer - nice package
# :: means give me a function from the package
# eval=False, tidy=TRUE - the code is being tidied, e.g. if don't put space between y and =, it does it in the output
```{r eval=FALSE, tidy=TRUE}
DataExplorer::create_report(iris_raw, y="Species")
```

##Querying rows
#pipelining - CTRL SHIFT M - %>% 
```{r}
iris_raw %>% 
  head(10) # values from top of data (top 10 rows). Sample of data
ttail(10) # gets the values from the bottom of data set (bottom 10 rows)

iris_raw %>% 
  sample_n(10) # show random ten rows + all columns

iris_raw %>% 
  sample_frac(0.10) # a fraction of data

# the lines before are equal to these:

head(iris_raw, 10)
tail(iris_raw, 10)
sample_n(iris_raw, 10)
sample_frac(iris_raw, 10)
```


##Where

```{r}
iris_raw %>% 
  filter(Species=="setosa", Sepal.Length >5.5)

filter(iris_raw, Species == "setosa", Sepal.Length >5.5)
```

# what does | mean - pipe

```{r}
iris_raw %>% 
  filter((Species=="setosa" | Sepal.Length > 5.5), Sepal.Width < 3)
```


```{r}
iris_raw %>%
  filter_if(is.numeric, ~ . < 5) #if too many columns. can only filter if value is numeric. Since we don't know which column title it would be then we use ~ and then . as place holder for title and then condition. Here, ~ is equivalent to "all variable" (all_vars)

iris_raw %>% 
  filter_if(is.numeric, any_vars( . <5)) # if any of the values in a row are TRUE to this condition, include it

iris_raw %>% 
  filter_if(is.numeric, all_vars( . <5)) # only if ALL values of a row are TRUE
```

## Filtering exercises 

# Exercise 1: Write a filter that gets all action movies from the movies dataset via the ggplot2movies package

```{r}
library(ggplot2movies)  # get library
movies # check the data frame
write_csv(movies, "movies.csv") # write the dataframe onto csv file
movies_data = read_csv ("movies.csv") # assign the read file to movies_data 
spec(movies_data) #to get more info on columns
head (movies_data) # get the top ten rows to get to know the data
movies_data %>% 
  sample_n(10) # see a random sample of data (10rows)
movies_data %>% filter(Action == "1") # filter the films which are classified as action by having 1 in the action column
col_double() #??
```
# Exercise 2: Write a filter that removes films lasting more than 6 hours from the movies dataset

```{r}
movies_data %>% 
  sample_n(10)
60*6
movies_data %>% 
  filter(length <= 360)
movies_data %>% 
  filter(length > - 360) # can you exclude values like this?? but then what if you have negative values? why is there more rows in this code than the previous?
```
## Exercise 3:[ADVANCED] Write a filter that checks to see if any of the films don’t have any genres flagged at all


```{r}
movies_data %>% 
  filter_at(vars(Action, Animation, Comedy, Drama, Documentary, Romance, Short), all_vars(. == 0))
```

library(beepr)
