---
title: "Data pipelines"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
write_csv(iris, "iris.csv")
```

## read

```{r}
read_csv("iris.csv")
iris_raw = read_csv("iris.csv")
```
## you can also read and write in excel
```{r}
library(readxl)
library(writexl)
write_xlsx(iris, "iris.xlsx")
read_excel("iris.xlsx")
```

##Exploratory analysis
Data explorer - nice package
# :: means give me a function from the package
# eval=False, tidy=TRUE - the code is being tidied, e.g. if don't put space between y and =, it does it in the output
```{r eval=FALSE, tidy=TRUE}
DataExplorer::create_report(iris_raw, y="Species")
```

##Querying rows
#pipelining - CTRL SHIFT M - %>% 
```{r}
iris_raw %>% 
  head(10) # values from top of data (top 10 rows). Sample of data
ttail(10) # gets the values from the bottom of data set (bottom 10 rows)

iris_raw %>% 
  sample_n(10) # show random ten rows + all columns

iris_raw %>% 
  sample_frac(0.10) # a fraction of data

# the lines before are equal to these:

head(iris_raw, 10)
tail(iris_raw, 10)
sample_n(iris_raw, 10)
sample_frac(iris_raw, 10)
```


##Where

```{r}
iris_raw %>% 
  filter(Species=="setosa", Sepal.Length >5.5)

filter(iris_raw, Species == "setosa", Sepal.Length >5.5)
```

# what does | mean - pipe

```{r}
iris_raw %>% 
  filter((Species=="setosa" | Sepal.Length > 5.5), Sepal.Width < 3)
```


```{r}
iris_raw %>%
  filter_if(is.numeric, ~ . < 5) #if too many columns. can only filter if value is numeric. Since we don't know which column title it would be then we use ~ and then . as place holder for title and then condition. Here, ~ is equivalent to "all variable" (all_vars)

iris_raw %>% 
  filter_if(is.numeric, any_vars( . <5)) # if any of the values in a row are TRUE to this condition, include it

iris_raw %>% 
  filter_if(is.numeric, all_vars( . <5)) # only if ALL values of a row are TRUE
```

## Filtering exercises 

# Exercise 1: Write a filter that gets all action movies from the movies dataset via the ggplot2movies package

```{r}
library(tidyverse)
library(ggplot2movies)  # get library
ggplot2movies::movies #look at waht data is in the package
movies # check the data frame
write_csv(movies, "movies.csv") # write the dataframe onto csv file
movies_data = read_csv ("movies.csv") # assign the read file to movies_data 
spec(movies_data) #to get more info on columns
head (movies_data) # get the top ten rows to get to know the data
movies_data %>% 
  sample_n(10) # see a random sample of data (10rows)
movies_data %>% filter(Action == "1") # filter the films which are classified as action by having 1 in the action column
col_double() #??
```
# Exercise 2: Write a filter that removes films lasting more than 6 hours from the movies dataset

```{r}
movies_data %>% 
  sample_n(10)
60*6
movies_data %>% 
  filter(length <= 360) %>%  
  summarise(max(length))

movies_data %>% 
  filter(length > - 360) # can you exclude values like this?? but then what if you have negative values? why is there more rows in this code than the previous? CANT DO THIS. The function only returns a TRUE or FALSE. SO need to make sure that the things we dont want return as false.

```
## Need to have the %>%  sign at the end of each line to carry on pipeline. Can save it in the end by assigning it to a new name with -> arrow
## Exercise 3:[ADVANCED] Write a filter that checks to see if any of the films don’t have any genres flagged at all


```{r}
movies_data %>% 
  filter_at(vars(Action, Animation, Comedy, Drama, Documentary, Romance, Short), all_vars(. == 0))

movies_data %>% 
  filter_at(vars(Action:Short), all_vars (. ==0))
```

library(beepr)

## Can use select function to select certain columns, or exclude

```{r}
iris %>% 
  select(Sepal.Length, Petal.Width)

iris %>% 
  select(-Sepal.Length)

iris %>% 
  select(-(Sepal.Length:Petal.Width))

iris %>% 
  select(starts_with("Sepal"))

iris %>% 
  select_if(is.numeric) %>% 
  prcomp()
```
##21.5 Exercises
# 1: Output the movies dataset with the column budget changed to budget_if_known

```{r}
movies %>% 
  rename(budget_if_known = budget) ->
  movies_cleaned

movies %>% 
  select(budget_if_known = budget, everything())
```
## Longer pipelines
Show sepal details for setosa flowers

```{r}
iris %>% 
  filter(Species == "setosa") %>% 
  select(starts_with("Sepal"))
```
## Changing or adding columns

```{r}
iris %>% 
  mutate(Sepal.Length = Sepal.Length - 0.1, 
         Sepal.Area = Sepal.Length * Sepal.Width) # creates a new column at the end and uses the corrected values.

iris %>% 
  mutate(Species_trun = str_trunc(as.character(Species),3), Diff.Sepal.Length = (Sepal.Length - mean(Sepal.Length))/mean(Sepal.Length))
```

## Exercise
# Add a column to movies that says how much the length differes from the median.

```{r}
movies %>% 
  mutate(diff_length = length - median(length)) ->
  movies_diff
movies_diff
```
## Aggregating our data
- groups = group by
- aggregation (how you are aggregating it) - another term is summarise

```{r}
iris %>% 
  group_by(Species) %>% 
  summarise(count = n(), avg.Sepal.Length = mean(Sepal.Length))

iris %>% 
  group_by(Species) %>% 
  summarise_all(mean)

iris %>% 
  mutate(Diff.Sepal.Length = (Sepal.Length - mean(Sepal.Length))/mean(Sepal.Length)) %>% # difference to total mean
  group_by(Species) %>% 
  mutate(Species.Diff.Sepal.Length = (Sepal.Length - mean(Sepal.Length))/mean(Sepal.Length)) # the difference is now calculated on a species basis ( diff of a value compared to the species mean)
```
## Section 26.7
# 1- Find the min and max values for length in movies
# 2- In a pipeline, find out what proportion of films are Action movies, Action Comedies, Comedies, and neither Action or Comedy

```{r}
# movie length range
movies %>% 
  summarise(min(length), max(length)/60)
#proportions
movies %>% 
  group_by(Action,Comedy) %>% 
  summarise(n=n()) %>% 
  mutate(per=n/sum(n)) %>%  # dont add up to 100 ...?
  ungroup() %>% 
  mutate(per=n/sum(n)) %>% 
  summarise(sum(per)) # check it adds up
```

